An Open-Source Implementation of the CMPS Algorithm for Assessing  Similarity of Bullet Signatures

### 2021-04-29 CMPSpaper_writing/CMPSpaper_writing.Rmd
couple of questions

Conceptually, the CMPS algorithm consists of \hh{three main} steps:

1. **cut the reference signature into consecutive, non-overlapping and equal-length basis segments:**
The command `get_segs(x, len=50)` implements this step:  it takes a bullet signature  in the format of a numeric vector and cuts it into consecutive, non-overlapping and equal-length segments of length `len`.

2. **register each basis segment on the comparison signature:** 
Calculate the cross-correlation curve between a basis segment and the comparison signature (using function `get_ccf4`).
A position is considered "good" \wj{if this position results in a peak in the cross-correlation between the basis segment and the comparison}
<!-- the cross-correlation between the basis segment and the comparison is above a pre-defined threshold.  -->
The best registration position is decided upon using a multi-peak inspection with or without different segment lengths:

    2.1. **multi-peak inspection (without different segment lengths):** each basis segment identifies positions of the \wj{top 5 highest peaks (if `npeaks.set = 5`) in the cross-correlation curve}. Small segment lengths might indicate false positive peaks. Therefore all 5 peaks are considered as candidates for the "best" registration position.

    2.2. **multi-peak inspection with different segment lengths:** for this version of the CMPS algorithm, each basis segment finds only one "best" registration position. This "best" registration position can be found by conducting multi-peak inspection multiple rounds. The first round is executed as stated in step 2.1. After that first round, the segment length is increased, a new cross-correlation curve is computed, and another set of  peaks in the cross-correlation curve is identified together with the candidate positions. Typically, fewer number of peaks are identified in this round (because the increased segment length reduces the chance of false positive peaks). This process can be repeated multiple times. In the last round, only one highest peak should be identified. If the same or 'similar enough' (see step 3) position can be identified for all different segment length, this position will be selected as the "best" registration position and is called the "consistent correlation peak" (ccp). \wj{In order to use this version of the CMPS algorithm, we set `npeaks.set` as a numeric vector of multiple values. For example, if we set `npeaks.set = c(5, 3, 1)`, 5 peaks will be identified in the first round, 3 peaks in the second round, and 1 peak in the final round.}

3. **determine the congruent registration position**:
candidate positions across different segments are compared using a small tolerance \wj{parameter `Tx`} and basis segments with a congruent registration position are determined. These segments are called the "congruent matching profile segments" (CMPS), and the number of CMPS is the CMPS score of the comparison.

@@

```{r, eval=FALSE}
seg.plot <- cmps_segment_plot(cmps_with_multi_scale, seg.idx = 6)
```

@@

```{r, eval=FALSE}
seg.plot[[1]]$segment_plot
```

```{r, echo=FALSE, eval=FALSE, out.width="300px", fig.cap="segment\\_plot for segment 6 at scale 1. The original position of segment 6 is indicated by the solid black line. Positions where the segment achieves the 5 highest cross-correlations are indicated by the dashed line segments. "}
knitr::include_graphics("img/segplot1.png", dpi = 100)
```

>>>>

Conceptually, the CMPS algorithm consists of three main steps:

1. **cut the reference signature into consecutive, non-overlapping and equal-length basis segments:**
The command `get_segs(x, len=50)` implements this step:  it takes  bullet signature `x` in the format of a numeric vector and cuts it into consecutive, non-overlapping and equal-length segments of length `len`.

2. **register each basis segment on the comparison signature:** 
Calculate the cross-correlation curve between a basis segment `x` and the comparison signature `y` using the function `get_ccf4(x, y, ...)`.
A position is considered "good" if it results in a peak in the cross-correlation between the basis segment and the comparison.
<!-- the cross-correlation between the basis segment and the comparison is above a pre-defined threshold.  -->
The best registration position is decided upon using a multi-peak inspection with or without different segment lengths:

    2.1. **multi-peak inspection (without different segment lengths):** each basis segment identifies positions of the top 5 highest peaks (if `npeaks.set = 5`) in the cross-correlation curve. Small segment lengths might indicate false positive peaks. Therefore all 5 peaks are considered as candidates for the "best" registration position.

    2.2. **multi-peak inspection with different segment lengths:** for this version of the CMPS algorithm, each basis segment finds only one "best" registration position. This "best" registration position can be found by conducting multi-peak inspection multiple rounds. The first round is executed as stated in step 2.1. After that first round, the segment length is increased, a new cross-correlation curve is computed, and another set of  peaks in the cross-correlation curve is identified together with the candidate positions. Typically, fewer number of peaks are identified in this round (because the increased segment length reduces the chance of false positive peaks). This process can be repeated multiple times. In the last round, only one highest peak should \hh{XXX should? this is ambiguous. What happens if no peak or multiple peaks are found?} be identified. If the same or 'similar enough' (see step 3) position can be identified for all different segment length, this position will be selected as the "best" registration position and is called the "consistent correlation peak" (ccp). \wj{In order to use this version of the CMPS algorithm, we set `npeaks.set` as a numeric vector of multiple values. For example, if we set `npeaks.set = c(5, 3, 1)`, 5 peaks will be identified in the first round, 3 peaks in the second round, and 1 peak in the final round.}

3. **determine the congruent registration position**:
candidate positions across different segments \hh{XXX how is this done in situation 2.1? } are compared using a small tolerance \wj{parameter `Tx`} and basis segments with a congruent registration position are determined. These segments are called the "congruent matching profile segments" (CMPS), and the number of CMPS is the CMPS score of the comparison.

@@

@@

\begin{figure}[htb]
```{r, opts.label="codefig", echo=TRUE, out.width='\\textwidth', warning = FALSE, fig.cap=" "}
#knitr::include_graphics("img/segplot1.png", dpi = 100)
seg.plot <- cmps_segment_plot(
  cmps_with_multi_scale, seg.idx = 6)
seg.plot[[1]]$segment_plot
```
\caption{\label{segplot} segment\_plot for segment 6 at scale 1. The original position of segment 6 is indicated by the solid black line. Positions where the segment achieves the 5 highest cross-correlations are indicated by the dashed line segments.}
\end{figure}
